---
title: "Chemokines and their receptors"
author: "original script: Perry Moerland\trefinement: Aldo Jongejan"
date: "November, 2023"
urlcolor: blue
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 5
    code_folding: show
---


```{r setup, include=FALSE,purl=FALSE}
knitr::opts_chunk$set(echo=TRUE,comment="#",message=FALSE,warning=FALSE,
                      fig.align='center',out.width="95%%",
                      fig.path = "21_chemokines_red_blue_AJ_20231130/",fig.show="hold")
```

```{r init}
# Set default for a good start 
source("initAnalysis.r")
#init()

outDir <- "21_chemokines_red_blue_AJ_20231130"
dir.create(outDir)

# Load packages
library(ggplot2)
library(multcomp)
library(RColorBrewer)
library(readxl)
library(Seurat)
library(GSEABase)
library(plyr)
library(dplyr)
```
  
This code is to be run on R < 4.3 and Seurat version < 4.4, so check:
  
```{r versions}
print(paste0("R version: ", R.version.string))
print(paste0("Seurat   : v", packageVersion("Seurat")))
```

First import the single-cell data:
```{r seurat}
# Read the 10X data
seurat <- readRDS("10_Reclustering_MP_and_MC/selectedClusters.subjectIntegrated.reclustered.rds")
seurat
DefaultAssay(seurat) <- "SCT"
Idents(seurat) <- "integrated_snn_res.1"
```
## With scale = TRUE

 Let's just retrieve two genesets also included in `15_pathway_visualization_20230419\enrichplot\pheatmap_AllComparisons_FDR.pdf`:
 
- [REACTOME_CHEMOKINE_RECEPTORS_BIND_CHEMOKINES](https://www.gsea-msigdb.org/gsea/msigdb/cards/REACTOME_CHEMOKINE_RECEPTORS_BIND_CHEMOKINES)
- [KEGG_CYTOKINE_CYTOKINE_RECEPTOR_INTERACTION](https://www.gsea-msigdb.org/gsea/msigdb/human/geneset/KEGG_CYTOKINE_CYTOKINE_RECEPTOR_INTERACTION.html)

and then visualize these. Note that these two genesets are large, especially
the one from KEGG.

```{r msigdb-sets}
## Import the MSigDB database v2023.1.Hs genesets
c2BroadSets = getGmt(con="../../../../../MSigDB/v2023.1.Hs/c2.all.v2023.1.Hs.symbols.gmt",geneIdType=SymbolIdentifier(), collectionType=BroadCollection(category="c2"))
BroadSets = geneIds(c2BroadSets)

# Select the two genesets
BroadSets <- BroadSets[grepl("REACTOME_CHEMOKINE_RECEPTORS_BIND_CHEMOKINES|KEGG_CYTOKINE_CYTOKINE_RECEPTOR_INTERACTION", names(BroadSets))]
sapply(BroadSets, length)
```
  
### First, only select the macrophage clusters

```{r macrophage_selection}
# Only include clusters 0, 1, 4 and 8
seurat.selClust <- subset(seurat, subset = integrated_snn_res.1 %in% c(0,1,4,8)) 
seurat.selClust@meta.data$integrated_snn_res.1 <- factor(seurat.selClust@meta.data$integrated_snn_res.1)
seurat.selClust
```
  
We are not using the REACTOME set here ...  
  
### KEGG


```{r kegg_in_seurat}
kegg <- BroadSets[[2]]
# These genes are not present in the SCT assay
setdiff(kegg, rownames(seurat.selClust))
```

Yosta is mentioning some genes in her paper.  

```{r gene_presence_kegg}
genesYosta <- c('CX3CR1', 'CCR1', 'CCR2','IL1B','TNF','CXCL2','CXCL3','CXCL8','IL18','IL12','IL15','CXCL16','CSF1R','CSF1','IFNGR2','VEGFA', 'VEGFB')
genesYosta %in% kegg
```

So, to begin with not all of them are present ... IL12    

#### FindAllMarkers - select based on avg_log2FC and take the top 15
  
Here, I select for an avg_log2FC > 0.2 and then take the top 15 per contrast

```{r Fig2D_kegg, dev=c('png','pdf','tiff'), fig.width=18}
kegg_sorted <- sort(intersect(kegg, rownames(seurat.selClust)))
length(kegg_sorted)

# In order to run FindAllMarkers, we need to run PrepSCTFindMarkers first
seurat.selClust <- PrepSCTFindMarkers(seurat.selClust)

# Lower the cut-offs in order to get most out of it
df.findAllMarkers <- FindAllMarkers(seurat.selClust, features=kegg_sorted, only.pos=TRUE, min.pct = 0.0, logfc.threshold = 0.0)

dim(df.findAllMarkers)  # There are 164 genes in the KEGG selection

# We miss some genes now, because there still is a filter on number of cells or something similar

library(dplyr)
df.findAllMarkers %>%
    group_by(cluster) %>%
    dplyr::filter(avg_log2FC > 0.2) %>%
    slice_head(n = 15) %>%
    ungroup() -> top15

c('CX3CR1', 'CCR1', 'CCR2','IL1B','TNF','CXCL2','CXCL3','CXCL8','IL18','IL12','IL15','CXCL16','CSF1R','CSF1','IFNGR2','VEGFA', 'VEGFB') %in% top15$gene

dim(top15)

DotPlot(seurat.selClust, features = unique(top15$gene), cols="RdBu", dot.scale = 9) + RotatedAxis() +
  theme(axis.text.x = element_text(size = 8)) 

# And save the data
write.table(df.findAllMarkers, paste(outDir,"kegg_findAllMarkers.txt", sep="/"), sep="\t", 
            quote = FALSE, row.names = TRUE, col.names = NA)

```
  
#### And extended figure 5

```{r ExtendedFig5_kegg, dev=c('png','pdf','tiff'), fig.width=18}
kegg <- BroadSets[[2]]
kegg <- sort(intersect(kegg, rownames(seurat)))
length(kegg)

head(kegg)

# Too large, so let's split it in half
kegg1 <- kegg[1:82]
kegg2 <- kegg[82:length(kegg)]

# The levels of the clusters are in order ... 
print(levels(seurat@meta.data$integrated_snn_res.1))

DotPlot(seurat, features = kegg1, cols="RdBu", dot.scale = 9) + RotatedAxis() +
  theme(axis.text.x = element_text(size = 8)) 

DotPlot(seurat, features = kegg2, cols="RdBu", dot.scale = 9) + RotatedAxis() +
  theme(axis.text.x = element_text(size = 8)) 

```

### Session Info
```{r sessionInfo}
print(sessionInfo())
```